---
title: "Logistic Regression"
author: "Joshua Harasaki, Tiffany Chiang, Katie Huynh"
date: "3/13/2022"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(tidyr)
library(ggplot2)
```

## Loading Data

```{r}
data = read.csv('./data/processed/processed_data.csv')  # load data
head(data)  # print first 5 rows of data
```

```{r}
data$Marital_Status <- as.factor(data$Marital_Status)
head(data)
```

```{r}
# create column telling us whether customer responded to any  
data$AcceptedCmp <- as.factor(ifelse(data$TotalAcceptedCmp > 0, 1, 0))
table(data$AcceptedCmp)
```

```{r}
qplot(AcceptedCmp, Total_Spending, data = data, geom = "boxplot")
```

```{r}
glm.fit = glm(AcceptedCmp ~ Education + Marital_Status + Income + Recency + NumWebVisitsMonth + 
                Total_Spending + Age + Children + TotalNumPurchases, data=data, family = binomial)
```

```{r}
# summarize the logistic regression model
summary(glm.fit)
```

```{r}
prob.training = predict(glm.fit, type = "response")
```

```{r}
pred_data = data %>% mutate(predAccept = as.factor(ifelse(prob.training <= 0.5, "No", "Yes")))
table(pred = pred_data$predAccept, true = pred_data$AcceptedCmp)
```

#### Interpreting Coefficients

-   From the summary of the fitted logistic regression model, we found that the variables Marital_Status, Income, Recency, NumWebVisitsMonth, Total_Spending, Children, and TotalNumPurchases are statistically significant.

**Holding all other variables fixed...**

-   For every dollar increase in salary, the log odds of accepting a promotion increases by 0.00001428.

-   The indicator function of "Married" has a regression coefficient -0.2928, meaning that a customer being married versus being single changes the log odds of accepting a promotion by -0.2928. Since exp(-.2928) = 0.74617, we can say the odds of accepting a promotion for a married customer is 0.74617 times greater than the odds of accepting a promotion for a single customer. In simplest terms, a customer being married indicates that there is a slightly less probability that that customer will accept a promotion offer.

-   For every one unit increase in recency, the log odds of accepting a promotion decreases by 0.0105.

-   For every one unit increases in the number of website visits in the last month, the log odds of accepting a promotion increases by 0.2584.

-   For every one unit increases in children, the log odds of accepting a promotion decreases by 0.2553.

-   For every one unit increase in the total number of purchases, the log odds of accepting a promotion decreases by 0.0365.

#### Findings from evaluating the coefficients of the logisitic regression model

From our exploratory data analysis, we found that total spending and total number of purchases are highly positively correlated. So, it was interesting to find that total spending has a positive correlation with a customer accepting a promotion, but total number of purchases has a negative correlation with a customer accepting a promotion.

Let's group customers who accepted a promotion and those who did not, to see how total number of purchases and total spending relates:

```{r}
# calculate the mean total number of purchases amoungst customers
plot_mean <- data %>%
  group_by(AcceptedCmp) %>%
  summarize(mean_num_purchases = round(mean(TotalNumPurchases), digits = 3))

# plot mean total number of purchases
ggplot(plot_mean, 
       aes(x = factor(AcceptedCmp,
                      labels = c("Did not accept", "Accepted")),
           y = mean_num_purchases)) +
  geom_bar(stat = "identity", fill = "cornflowerblue") +
  geom_text(aes(label = mean_num_purchases, vjust = -0.25)) +
  labs(x = "Customer Response", y = "Mean Total Number of Purchases") 
```

```{r}
# calculate the mean total spending amoungst customers
plot_mean <- data %>%
  group_by(AcceptedCmp) %>%
  summarize(mean_total_spending = round(mean(Total_Spending), digits = 3))

# plot mean total number of purchases
ggplot(plot_mean, 
       aes(x = factor(AcceptedCmp,
                      labels = c("Did not accept", "Accepted")),
           y = mean_total_spending)) +
  geom_bar(stat = "identity", fill = "cornflowerblue") +
  geom_text(aes(label = mean_total_spending, vjust = -0.25)) +
  labs(x = "Customer Response", y = "Mean Total Spending") 
```

```{r}
mean(data$Total_Spending)
```

As you can see in the chart, we see that the mean of the total number of purchases is higher for customers who don't accept campaigns. This may be for a variety of reasons, but we hypothesize that the promotions may be for more expensive items, thus explaining why although a customer spending more
